---
title: "Untitled"
author: "Kwame Forbes"
date: "11/2/2022"
output: html_document
---

loading Dependencies
```{r}
library(DESeq2)
source("/pine/scr/k/w/kwamek/pengda_collab/IRFinder-2/bin/DESeq2Constructor.R")  #Load IRFinder-related function
```

```{r}
results = read.table("results.txt")
paths = as.vector(results$V1)                                           
experiment = read.table("Experiment.txt",header=T)                       
experiment$Condition=factor(experiment$Condition,levels=c("WT","T200M"))    
rownames(experiment)=NULL                                               
```
# WARNING: make sure the rownames of `experiment` is set to NULL. 
# WARNING: users MUST check if the order of files in the `path` matches the order of samples in `experiment` before continue  


```{r}
metaList=DESeqDataSetFromIRFinder(filePaths=paths, designMatrix=experiment, designFormula=~1)
```
# The above line generate a meta list contains four slots
# First slot is a DESeq2 Object that can be directly pass to DESeq2 analysis.  
# Second slot is a matrix for trimmed mean of intron depth
# Third slot  is a matrix for correct splicing depth flanking introns
# Fourth slot is a matrix for maximum splicing reads at either ends of introns
# We build a “null” regression model on the interception only. 
# A “real” model can be assigned either here directly, or in the downstream step. See below


```{r}
dds = metaList$DESeq2Object                       # Extract DESeq2 Object with normalization factors ready
colData(dds)                                      # Check design of matrix
```

# Please note that sample size has been doubled and one additional column "IRFinder" has been added.
# This is because IRFinder considers each sample has two sets of counts: one for reads inside intronic region and one for reads at splice site, indicating by "IR" and "Splice" respectively.
# "IRFinder" is considered as an additional variable in the GLM model.
# Please also be aware that size factors have been set to 1 for all samples. Re-estimation of size factors is NOT recommended and is going to bias the result.
# More details at the end of the instruction.
```{r}
design(dds) = ~Condition + Condition:IRFinder     # Build a formula of GLM. Read below for more details. 
dds = DESeq(dds)                                  # Estimate parameters and fit to model
resultsNames(dds)                                 # Check the actual variable name assigned by DESeq2
```
 

```{r}
res.WT = results(dds, name = "ConditionWT.IRFinderIR")
```
# This tests if the number of IR reads are significantly different from normal spliced reads, in the WT samples.
# We might only be interested in the "log2FoldChange" column, instead of the significance.
# This is because "log2FoldChange" represents log2(number of intronic reads/number of normal spliced reads).
# So we the value of (intronic reads/normal spliced reads) by


```{r}    
WT.IR_vs_Splice=2^res.WT$log2FoldChange
```

# As IR ratio is calculated as (intronic reads/(intronic reads+normal spliced reads))
# We can easily convert the above value to IR ratio by
```{r}    
IRratio.WT = WT.IR_vs_Splice/(1+WT.IR_vs_Splice)
    
# Similarly, we can get IR ratio in the T200M samples
res.T200M = results(dds, name = "ConditionT200M.IRFinderIR")
T200M.IR_vs_Splice=2^res.T200M$log2FoldChange
IRratio.T200M = T200M.IR_vs_Splice/(1+T200M.IR_vs_Splice)

# Finally we can test the difference of (intronic reads/normal spliced reads) ratio between WT and T200M
res.diff = results(dds, contrast=list("ConditionT200M.IRFinderIR","ConditionWT.IRFinderIR"))  
```   
# We can plot the changes of IR ratio with p values
# In this example we defined significant IR changes as
# 1) IR changes no less than 10% (both direction) and 
# 2) with adjusted p values less than 0.05
```{r}
IR.change = IRratio.T200M - IRratio.WT
plot(IR.change,col=ifelse(res.diff$padj < 0.05 & abs(IR.change)>=0.1, "red", "black"))
```
# PLEASE NOTE    
# You might see dots at the same horizontal level (same IR change) either marked as significant (red) or not (black)
# This is because the Wald test applied above is testing on fold changes instead of absolute changes
# For example, both changes from 0.01 to 0.11 and from 0.8 to 0.9 have absolute changes of 0.1.
# But the fold changes for them are 11 folds and 1.1 folds respectively, and lead to different p values