---
title: "IRFinder_Pendga"
author: "Kwame Forbes"
date: "10/6/2022"
output: html_document
---
Load Dependencies
```{r}
library(ggplot2)
library(org.Hs.eg.db)
```
variables
```{r}
LFC = 1.0
PADJ = 0.05
```



loading comparing WT-MUT
investigate differential IR between two conditions with replicates 
Minimum IR ratio of  0.05 met in at least one sample to consider the intron.
Discard the LowCover and LowSplicing 
```{r}
MUT_WT = read.csv("/pine/scr/k/w/kwamek/pengda_collab/irfiles_Pengda/WT-MUT-Analysis_2/MUT_WT_DESeq2.tsv",sep = "\t")
MUT_WT$ensem = gsub("\\..*","",MUT_WT$ID)
MUT_WT$symbol = mapIds(org.Hs.eg.db,
                           keys = MUT_WT$ensem,
                           column = "SYMBOL",
                           keytype="ENSEMBL",
                           multiVals = "first")
AS = as.data.frame(rep("WTvT200M",dim(MUT_WT)[1]))
names(AS)[names(AS) == 'rep("WTvT200M", dim(MUT_WT)[1])'] <- "Comparison"
MUT_WT = cbind(MUT_WT,AS)
MUT_WT$Comparison <- as.factor(MUT_WT$Comparison)
MUT_WT$Event <- ifelse(abs(MUT_WT$log2FoldChange) >= LFC & MUT_WT$padj <= PADJ,"Signifcant","Not Significant")
write.csv(MUT_WT,"WTvsT200M_IRFinder.csv")
MUT_WT_sig <- subset(MUT_WT,Event =="Signifcant")

```



plot WT-MUT
```{r}
p <- ggplot(data = MUT_WT,mapping= 
              aes(x =log2FoldChange,y =-log10(padj),color=Event))+ theme_classic() + geom_point() + 
   ggtitle("WT vs. T200M Differential Intron Retention") +xlab("Log2FoldChange") + ylab("-log10(padj)") + scale_color_manual(values=c("black", "red"))
p <- p + theme(plot.title = element_text(color="black", size=21, face="bold.italic"),
         axis.title.x = element_text(color="black", size=20, face="bold"),
         axis.title.y = element_text(color="black", size=20, face="bold"),
         axis.text = element_text(size = 21),
         legend.text = element_text(size=17),
         legend.title = element_text(size=20))
p
```



loading comparing WT-EV
investigate differential IR between two conditions with replicates 
Minimum IR ratio of  0.05 met in at least one sample to consider the intron.
Discard the LowCover and LowSplicing 
```{r}
EV_WT = read.csv("/pine/scr/k/w/kwamek/pengda_collab/irfiles_Pengda/WT-EV-Analysis_2/EV_WT_DESeq2.tsv",sep = "\t")
EV_WT$ensem = gsub("\\..*","",EV_WT$ID)
EV_WT$symbol = mapIds(org.Hs.eg.db,
                           keys = EV_WT$ensem,
                           column = "SYMBOL",
                           keytype="ENSEMBL",
                           multiVals = "first")
AS = as.data.frame(rep("WTvEV",dim(EV_WT)[1]))
names(AS)[names(AS) == 'rep("WTvEV", dim(EV_WT)[1])'] <- "Comparison"
EV_WT = cbind(EV_WT,AS)
EV_WT$Comparison <- as.factor(EV_WT$Comparison)
EV_WT$Event <- ifelse(abs(EV_WT$log2FoldChange) >= LFC & EV_WT$padj <= PADJ,"Signifcant","Not Significant")
write.csv(EV_WT,"WTvEV_IRFinder.csv")
EV_WT_sig <- subset(EV_WT,Event =="Signifcant")
```
plot EV-WT
```{r}
r <- ggplot(data = EV_WT,mapping= 
              aes(x =log2FoldChange,y =-log10(padj),color=Event))+ theme_classic() + geom_point() + 
   ggtitle("WT vs EV Differential Intron Retention") +xlab("Log2FoldChange") + ylab("-log10(padj)") + scale_color_manual(values=c("black", "red"))
r <- r + theme(plot.title = element_text(color="black", size=21, face="bold.italic"),
         axis.title.x = element_text(color="black", size=20, face="bold"),
         axis.title.y = element_text(color="black", size=20, face="bold"),
         axis.text = element_text(size = 21),
         legend.text = element_text(size=17),
         legend.title = element_text(size=20)) 
r
```




loading comparing MUT-EV
investigate differential IR between two conditions with replicates 
Minimum IR ratio of  0.05 met in at least one sample to consider the intron.
Discard the LowCover and LowSplicing 
```{r}
EV_MUT = read.csv("/pine/scr/k/w/kwamek/pengda_collab/irfiles_Pengda/T200M-EV-Analysis_2/EV_MUT_DESeq2.tsv",sep = "\t")
EV_MUT$ensem = gsub("\\..*","",EV_MUT$ID)
EV_MUT$symbol = mapIds(org.Hs.eg.db,
                           keys = EV_MUT$ensem,
                           column = "SYMBOL",
                           keytype="ENSEMBL",
                           multiVals = "first")
AS = as.data.frame(rep("T200MvEV",dim(EV_MUT)[1]))
names(AS)[names(AS) == 'rep("T200MvEV", dim(EV_MUT)[1])'] <- "Comparison"
EV_MUT = cbind(EV_MUT,AS)
EV_MUT$Comparison <- as.factor(EV_MUT$Comparison)
EV_MUT$Event <- ifelse(abs(EV_MUT$log2FoldChange) >= LFC & EV_MUT$padj <= PADJ,"Signifcant","Not Significant")
write.csv(EV_MUT,"T200MvEV_IRFinder.csv")
EV_MUT_sig <- subset(EV_MUT,Event =="Signifcant")
```


plot EV-MUT
```{r}
s <- ggplot(data = EV_MUT,mapping= 
              aes(x =log2FoldChange,y =-log10(padj),color=Event))+ theme_classic() + geom_point() + 
   ggtitle("T200M vs. EV Differential Intron Retention") +xlab("Log2FoldChange") + ylab("-log10(padj)") + scale_color_manual(values=c("black", "red"))
s <- s + theme(plot.title = element_text(color="black", size=21, face="bold.italic"),
         axis.title.x = element_text(color="black", size=20, face="bold"),
         axis.title.y = element_text(color="black", size=20, face="bold"),
         axis.text = element_text(size = 21),
         legend.text = element_text(size=17),
         legend.title = element_text(size=20)) 
s
```











Violin/boxplot for Comparisons
```{r}

big_ir <- rbind(EV_MUT,EV_WT,MUT_WT)
big_ir_sig <- subset(big_ir,Event == "Signifcant")

q <- ggplot(big_ir_sig, aes(x=Comparison, y=log2FoldChange,color =Comparison)) + geom_violin()

q <- q + stat_summary(fun =median, geom="point", shape=23, size=4)  + theme(plot.title = element_text(color="black", size=21, face="bold.italic"),
         axis.title.x = element_text(color="black", size=20, face="bold"),
         axis.title.y = element_text(color="black", size=20, face="bold"),
         axis.text = element_text(size = 21),
         legend.text = element_text(size=17),
         legend.title = element_text(size=20)) 

```


```{r}
wtvmut_as <- read.csv("/nas/longleaf/home/kwamek/AS_Pengda/WTvsT200M-50counts-filtered.csv",sep = ",")
wtvmut_ri <- subset(wtvmut_as,events =="RI")

sig_wtvmut <- subset(MUT_WT, log2FoldChange>=LFC&padj<=PADJ)

```




